---
title: "Week 1 – Intro"
author: "Erika Lee"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##updated R studio to version 4.3.3
#set up packages
install.packages(c("tidyverse", "dataRetrieval", "sf", "stars", "terra"))

install.packages(c("xgboost", "randomForest"))

library(sf)
library(stars)
library(terra)
library(tidyverse)
library(dataRetrieval)
```

## Class Notes

### Why public datasets?

-   cheaper way to test initial ideas

-   Expand local study or experiment to other ecosystems

### Tidying Messy Public data

-   Use R to do this, they have great packages

### Downloading Data

-   dataRetrieval package makes it easy to download data

-   Steps:

    -   Set up instructional tibbles (tables) that help understand what you will be doing later on

    -   Table acts as an instructional set for later

    -   Create parameter names as well for variables, then compile into a single list, then name list

    -   THIS is all just instructions, to run a function

-   Then you can run a loop that runs each parameter through the function to make a list of all the things we need - final list will be filled in with all dataframes that you want

-   Then you need to relate the lists (like a sheet in excel) together -\> bind all dataframes into a single dataframe through conc.long format

    example of code: conc.long \<- map_dfr(conc.list,rbind)

-   Then also need to get site information for each location

-   Save the data, so you can be able to load it easily

### How to Tidy the Data that's been saved

-   Convert data into readable information in a dataframe (shorter words you understand)

-   Make sure units are all the same, if not, convert into the same units

-   Make sure data is only showing the mean (or median, depending on what you're doing) of concentration samples - **you need to know time interval for samples!**

-   Then can convert into a st_as_sf to visualize data on map - making it into a spatial object

### 
